\documentclass{tufte-handout}
\usepackage{amsmath,amsthm}

\usepackage{pgfplots}
\pgfplotsset{width=\textwidth}

\newtheorem{claim}{Claim}[section]
\title{\sf Exact Algorithm for Independent Set}
\author{}

\newcommand{\mis}{\textsc{MIS}}
\newcommand{\nodeu}{\ensuremath{\tiny u}}
\newcommand{\nodev}{\ensuremath{\tiny v}}

\tikzset{
  auto,node distance=5mm,thick,
  remaining node/.style={thick,circle,draw,minimum size=3mm,fill=black!10},
  remaining edge/.style={thick},
  mis node/.style={dotted,circle,draw,minimum size=3mm,fill=blue!20},
  removed node/.style={dotted,circle,draw,minimum size=3mm},
  removed edge/.style={thin,dotted}
}

\begin{document}
\maketitle

\section{Independent Set Lab Report}
by Erik Westrup and Dmitry Basavin

\subsection{Correctness}
Algorithm R1 correctly computes $\alpha(G)$ because putting a node of degree 1 in the MIS is no worse than putting his neighbor in the MIS. This is because the neighbours could have more than one neighbour, thus putting it in the MIS could potentially block more verticies from being placed in the MIS.

\noindent
Algorithm R2 correctly computes $\alpha(G)$. In the first case when the cilque of size 3 is detected, then there is no way of picking a MIS greater than 1. Thus one of the vertices are accounted for and we recurse on G without this clique.

In the second case in the original graph G there are two options: 1) pick v, or 2) pick u and w. This corresponds, in the modified graph, to 1) picking all the grandchildren of v in G, and  2) picking z in the modified graph. This we can return 1 + recursion on the modified graph.

\subsection{Empirical Running time}

\paragraph{Experiments.}

\medskip
\noindent
\input{output.data}
    

The running times of algorithm (measured in the number of recursions, assuming $o(c^n)$, we take the logarithm of the number of recursive calls and fit a polynomial function of degree 1 to it. The exponent of the coefficient for the fitted line is $c$. ) ~$R_0$, $R_1$, and $R_2$ appear to be
$O(1.297^n)$ , $O(1.166^n)$, and $O(1.149^n)$, respectively..

\subsection{Theoretical Upper Bound}

Denote be $T_i(n)$ the worst runtime of algorithm Ri on \emph{any} graph on $n$ vertices.
Note that $T_i(n)$ is a non-decreasing function of $n$.
For $R_0$ we can conclude that
\begin{align*}
T_0(n) &\leq\max(T_0(n-1), T_0(n-1)+T_0(n-1-d_{\mbox{max}})) \\ &\leq T_0(n-1)+T_0(n-2)
\end{align*}
with $d_{\mbox{max}}$ the degree of the vertex we branch on. The hard part is the one when there are no isolated vertices, in which case the vertex $u$ we are branching on has at least one neighbor. 

For R1 we have that
\begin{align*}
	T_1(n) &\leq\max(T_1(n-1), T_1(n-2), T_1(n-1)+T_1(n-1-d_{\mbox{max}})) \\ &\leq T_1(n-1)+T_1(n-3)
\end{align*}

because we remove two nodes in the new case, and the last case always have $d_{max} \geq 2$.

For R2 we have that
\begin{align*}
	T_2(n) &\leq\max(T_2(n-1), T_2(n-2), T_2(n-2), T_2(n-3), T_2(n-1)+T_2(n-1-d_{\mbox{max}})) \\ &\leq T_2(n-1)+T_2(n-4)
\end{align*}
because we remove either 3 or 2  nodes in the new case, and the last case always have $d_{max} \geq 3$.


\paragraph{Worst Case Upper Bound}
The running times of algorithm~$R_0$, $R_1$, and $R_2$ are in
$O((\frac{1+\sqrt(5)}{2})^n)=O(1.618^n)$ , $O(1.465^n)$, and $O(1.38^n)$, respectively. \newpage

\section{Optional}
Add more ``algorithmic intelligence'' to the algorithm $R_2$ in order to tackle the instance in data/g130.in.
Try to make it run in less than 10,000,000 recursive calls. 

Suggestions for possible speed-ups:
\begin{itemize}
\item Is there some clever (branching) rule for vertices of degree 3?
\item Can we use the information of the largest found independent set so far, to reduce further computation time?
\item What if the graph gets disconnected?
\end{itemize}


\end{document}
